---
title: 'Course Project: Activity Monitoring Devices Data Analysis'
author: "SG"
date: "23/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "asis", message = FALSE, warning=FALSE)
```
## Data reading and processing
1. Code for reading in the dataset and/or processing the data
Packages "lubridate"  and "dplyr" are useful in analysing data related to dates and data frames, respectively.

```{r}
library(dplyr)
library(lubridate)
setwd("~/Online_Courses/Data Analysis R Foundations/Reproducible research/Project 1/RepData_PeerAssessment1")
data <- read.table(unz("activity.zip", "activity.csv"), sep=",", header = TRUE)
data$date <- as.Date(data$date)
```

Next, we perform data analysis as per the assignment's instructions. 

## Steps per day
2. Histogram of the total number of steps taken each day

```{r fig.align = 'center'}
step_day <- data %>% group_by(date) %>% summarize(stepsperday = sum(steps, na.rm = TRUE))
with(step_day, hist(stepsperday, breaks = 50, xlab = "Steps per day", ylab = "Frequency - Number of days", main = "Histogram of steps per day"))
```

3. Mean and median of steps taken each day
```{r}
mean <- mean(step_day$stepsperday)
median <- median(step_day$stepsperday)
```
The mean value is `r round(mean, 2)` and median value is `r median`.


